---
title: 'Lab 4: Does Prenatal Care Improve Infant Health?'
subtitle: 'W203: Statistics for Data Science'
date: "December 8, 2016"
author: 'Dan Guevarra, Bas Hendri, Saurabh Jaju ({beatspace9,dip,saurabh.jaju}@berkeley.edu)'
output:
  html_document: default
  pdf_document: default
---
```{r,echo=F,results='hide'}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
# http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/
# CC BY-SA
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

```{r}
load("bwght_w203.RData")
desc
```
```{r,warning=F,message=F}
library(car)
library(sandwich)
##  Hlavac, Marek (2015). stargazer: Well-Formatted Regression and Summary Statistics Tables.
library(stargazer)
library(reshape2)
library(ggplot2)
library(data.table)
```

<!--
# Assignment

Prepare a report addressing the question of whether prenatal care improves newborn health outcomes.
-->

## 1  Introduction
<!--
1. A brief introduction
-->

According to the National Institutes of Health, "prenatal care should begin as soon as a woman knows or suspects she is pregnant... prenatal care is important to help promote a healthy pregnancy. Women who do not seek prenatal care are three times as likely to deliver a low birth weight infant, and lack of prenatal care can also increase the risk of infant death." The following statistical analysis aims to substantiate these claims using a classical linear regression model.

https://www.nichd.nih.gov/health/topics/preconceptioncare/conditioninfo/Pages/default.aspx

## 2  Preliminary Explorations
<!--
2. A model building process, supported by exploratory analysis.  Your EDA should be interspersed with, and support, your modeling decisions.  In particular, you should use exploratory techniques to address
    - What transformations to apply to variables and what new variables should be created.
    - What variables should be included in each model
    - Whether model assumptions are met 
-->

```{r}
# race
race <- c(1,2,NA,3)
ilbw <- c(1,2,3)
data$mrace <- (race[apply(t(t(data[,c('mwhte','mblck','moth')])*c(1,2,4)),1,sum)])
data$frace <- (race[apply(t(t(data[,c('fwhte','fblck','foth')])*c(1,2,4)),1,sum)])
data$ilbw <- (ilbw[1+apply(data[,c('lbw','vlbw')],1,sum)])
is.factor.var <- c(14,24:26)
is.value.var <- c(1:11,21:23)
```

```{r}
# non-indicator vars:
ggplot(melt(data[,is.value.var]), aes(x=value)) +
    geom_histogram(bins=30) +
    facet_wrap(~variable, scales='free') +
    ggtitle('Histograms of non-indicator variables')


# indicator vars
ggplot(melt(data[,is.factor.var]), aes(x=factor(value))) +
    stat_count() +
    facet_wrap(~variable, scales='free') +
    ggtitle('Count of indicator variables')
```
<!-- suppress loess warnings -->
```{r,warning=F}
# warnings for loess due to discretized variables have been suppressed
scatterplotMatrix(data[,c('mage', 'meduc', 'monpre', 'npvis', 'fage', 'feduc', 'bwght', 'fmaps', 'omaps')])
```
```{r}
bwdt<-as.data.table(data)
bwdt[order(fmaps),.N,by=fmaps]
bwdt[order(cigs),.N,by=cigs]

```


- What transformations to apply to variables and what new variables should be created.

The non-indicator variables don't appear particularly conducive to log transformation. The as-supplied transformed variables `magesq` and `npvissq` appear more positively skewed their originals. The `lbwght` variable is also more negatively skewed than the non-transformed `bwght`. The education variables `meduc` and `feduc` have spikes at the end of traditional highschool and 4-year college educations. Indicator variables for degree of education may be a more appropriate grouping that weighs higher education separately.

Among the indicator variables, there are far more "normal" birthweights than low or very low birthweights, and more white mothers and fathers than black or other.

We create variables `mrace` and `frace` to summarize the races.

```{r,results='hide'}
bwdt[,mrace:=factor(mwhte*1+mblck*2+moth*3, 
                    levels=c(1, 2, 3),
                    labels=c('white', 'black', 'other'))]
bwdt[,frace:=factor(fwhte*1+fblck*2+foth*3, 
                    levels=c(1, 2, 3),
                    labels=c('white', 'black', 'other'))]
```
```{r,fig.width=12,fig.height=8,fig.align='center'}
ggplot(melt(bwdt[,.(fmaps,npvis,meduc,npvissq,feduc,mage)], id.vars='fmaps'), aes(x=value, y=fmaps)) +
    geom_jitter(alpha=I(1/20)) +
    facet_wrap(~variable, scales='free', ncol=3)
```
```{r,fig.width=12,fig.align='center'}
ggplot(melt(bwdt[,.(npvis,feduc,meduc)], id.vars='npvis'), aes(x=value, y=npvis)) +
    geom_jitter(alpha=I(1/20)) +
    facet_grid(.~variable, scales='free')
```

```{r,fig.width=12,fig.align='center'}
ggplot(melt(bwdt[,.(bwght,npvis,mage,feduc,meduc)], id.vars='bwght'), aes(x=value, y=bwght)) +
    geom_jitter(alpha=I(1/20)) +
    facet_grid(.~variable, scales='free')
```


- What variables should be included in each model

APGAR score is an aggregate rating of heart rate, respiratory effort, muscle tone, response to stimulation, and skin coloration. The one minute score is used to determine immediate medical response, but a low score doesn't directly translate to an unhealthy baby (premature or c-section births typically score low). The five minute score is a measure of progress and response to medical intervention and is a better indicator of a baby's overall health than the one minute score. The `fmaps` variable seems to be the appropriate dependent variable that describes infant health. However, the distribution of `fmaps` data is poorly suited for regression exercise as the variable is ordinal and an overwhelming majority (close to 90%) is concentrated in `fmaps==9`.

We opt for birthweight instead as the distribution is more well-behaved. Our models will not control for "low" and "very low" birthweight indicator variables. As a proxy for infant health, the ceteris paribus effect of `npvis` on birthweight will be evaluated against the "low" and "very low" birthweight thresholds of 2500 and 1500 grams respectively.


```{r}

```

- Whether model assumptions are met

## 3  Models
<!--
3. A minimum of three model speciï¬cations. In particular, you should include
    - One model with only the explanatory variables of key interest.
    - One model that includes only covariates that you believe increase the accuracy of your results without introducing bias.
    - One model that includes the previous covariates, but also covariates that may be problematic for one reason or another. 
-->

### 3.1 Model with only the explanatory variables of key interest

We are interested in discovering the effect of prenatal care, so for key interest variables we take count of prenatal visits `npvis` and months of prenatal care `monpre`. We additionally take `npvissq` to model diminishing return.

```{r}
(mod1b<-lm(bwght ~ npvis + npvissq + monpre, data=bwdt))
```

### 3.2 Model with only covariates without bias

We additionally take average cigarette smoked `cigs`, and add mother's age and squared age `mage` and `magesq` for dimishing return.

```{r}
(mod2b<-lm(bwght ~ npvis + npvissq + monpre + cigs + mage + magesq, data=bwdt))
```

### 3.3 Model with potentially problematic covariates

We also add average drink, and whether baby is male.

```{r}
(mod3b<-lm(bwght ~ npvis + npvissq + monpre + cigs + mage + magesq + drink + frace*mrace, data=bwdt))
```

## 4  Diagnostic Assessments
<!--
4. For your first model, a detailed assessment of the 6 CLM assumptions.  For additional models, you should check all assumptions, but only highlight major differences from your first model in your report.
-->

### 4.1 Model 1

#### MLR.1: Linear in Parameters

```{r,fig.width=12,fig.height=4,fig.align='center'}
ggplot(melt(bwdt[,.(bwght,npvis,npvissq,monpre)], id.vars='bwght'), aes(x=value, y=bwght)) +
    geom_jitter(alpha=I(1/20)) +
    facet_grid(.~variable, scales='free')
```

From the scatterplot matrix, we observe a major cluster of bivariate Gaussian distribution on top left for all predictors. However, in all cases there are also distribution of linearly increasing effect of predictor on birth weight.

#### MLR.2: Random Sampling

```{r}
ggplot(melt(bwdt[,.(bwght,npvis,npvissq,monpre)]), aes(x=value)) +
    geom_histogram(bins=30) +
    facet_wrap(~variable, scales='free') +
    ggtitle('Histograms of non-indicator variables')
```

From the histograms we observe that the sample has somewhat normally distributed variable values over the ranges.

#### MLR.3: No Perfect Collinearity

```{r,fig.width=8,fig.height=8,fig.align='center'}
par.orig <- par(no.readonly=T); par(mfrow=c(2,2)) # plot on 2x2

scatter.smooth(jitter(bwdt$npvis), jitter(bwdt$monpre), pch=20, col=rgb(0,0,0,0.05))
scatter.smooth(jitter(bwdt$npvissq), jitter(bwdt$monpre), pch=20, col=rgb(0,0,0,0.05))
scatter.smooth(jitter(bwdt$npvis), jitter(bwdt$npvissq), pch=20, col=rgb(0,0,0,0.05))
```

#### MLR.4: Zero Conditional Mean
```{r,fig.width=12,fig.height=12}
par(mfrow=c(2,2)) # plot diagnostics on 2x2

plot(mod1b)
```

#### MLR.5: Homoskedasticity


#### MLR.6: Normality



### 4.1 Model 2

#### MLR.1: Linear in Parameters
No changes effect MLR.1, as defined by the our linear model construction.

#### MLR.2: Random Sampling
```{r}
ggplot(melt(bwdt[,.(cigs,mage,magesq)]), aes(x=value)) +
    geom_histogram(bins=30) +
    facet_wrap(~variable, scales='free') +
    ggtitle('Histograms of non-indicator variables')

bwdt[order(cigs),.N,by=cigs]
bwdt[order(cigs),.N,by=cigs>0]
```

Histograms of the `mage` and `magesq` variables look approximately normal. The `cigs` variable is mostly 0, with a marginal number of positive cases. Omitting $NA$ cases of `cigs` shows about ~10% of mothers smoked at least 1 cigarette per day.

#### MLR.3: No Perfect Collinearity
```{r}
cor(bwdt[,.(bwght,npvis,npvissq,monpre,cigs,mage,magesq)], use='pairwise.complete.obs')
```

Aside from the expected correlation between `mage` and `magesq`, `npvis` and `npvissq`, there's no evidence of strong multicollinearity with the addition of `cigs`, `mage`, and `magesq`.

#### MLR.4: Zero Conditional Mean
```{r,fig.width=12,fig.height=12}
par(mfrow=c(2,2)) # plot diagnostics on 2x2

plot(mod2b)
```

There's no major difference in the residuals vs fitted values plot compared to model 1. The smoothing line still shows an approximate distribution of residuals about zero, with the same small grouping of values toward -3000.

#### MLR.5: Homoskedasticity
```{r}
bptest(mod2b)
```

The scale-location plot also looks similar to model 1. The smoothing line is fairly consistent with a minor uptick at higher fitted values (x-axis). The Breusch-Pagan test results in p<0.05, rejecting the null hypothesis (homoskedasticity), however the strong statistical significance is likely due to n=1832 observations.


#### MLR.6: Normality
```{r}
shapiro.test(mod2b$residuals)
```

The normal q-q plot shows deviation from normal at lower theoretical quantiles, and similar to the breusch-pagan test, the high number of observations influences the shapiro-wilks test and rejects normality with very high significance. No changes from model 1 and n=1832 invokes CLT.

### 4.1 Model 3

#### MLR.1: Linear in Parameters
#### MLR.2: Random Sampling
#### MLR.3: No Perfect Collinearity
#### MLR.4: Zero Conditional Mean
```{r,fig.width=12,fig.height=12}
par(mfrow=c(2,2)) # plot diagnostics on 2x2

plot(mod3b)
```

#### MLR.5: Homoskedasticity
#### MLR.6: Normality



## 5  Model Summary
<!--
5. A well-formatted regression table summarizing your model results.  Make sure that standard errors presented in this table are valid.  Also be sure to comment on both statistical and practical significance.
-->

```{r}
se.mod1b = sqrt(diag(vcovHC(mod1b)))
se.mod2b = sqrt(diag(vcovHC(mod2b)))
se.mod3b = sqrt(diag(vcovHC(mod3b)))

stargazer(mod1b, mod2b, mod3b, type = "text",
          se = list(se.mod1b, se.mod2b, se.mod3b),
          star.cutoffs = c(0.05, 0.01, 0.001))

(AIC(mod1b)-AIC(mod2b))/AIC(mod1b)
(AIC(mod2b)-AIC(mod3b))/AIC(mod2b)
```

Using heteroskedasticity-robust standard errors, we find p<0.05 statistical significance across all parameters in model 1, and all but `npvissq` coefficient in model 2. The addition of `drink` and parent race variables prevents vcovHC() calculation of standard errors for model 3 so we cannot comment on statistical signifiance, other than the relative drop in AIC from model 1 to 2 was 6.2%, whereas from model 2 to 3 was only 0.5%.

In terms of practical signifiance, model 1 shows that the `npvis` parameter has the greatest effect on birthweight (50 grams per prenatal care visit, ignoring the quadratic term for simplicity). The addition of `mage` and `cigs` in model 2 has a drastic effect on both the `npvis` parameter (50 drops to 40 grams per visit) and the intercept term ~2900 grams falls to ~1800 grams with zero prenatal vists. In addition, model 2 shows mother's age `mage` having a stronger effect than `npvis`. Proper ceteris paribus interpretation requires solving for the quadratic expression, which will be covered in the causal interpretation. The negative value `cigs` parameter agrees with the expected negative impact on infant health.

The addition of `drink` and race variables has marginal effect on the `npvis`, `mage`, and `cigs` parameters, which increase in magnitude slightly. The greater effect on practical significance occurs for the race-modification of the intercept term, which by itself corresponds with a white mother + white father. Notable race combinations are white mother + black father and "other" mother + "other" father, which both lower the intercept term (=1733-354) to less than 1500 grams, the very low birthweight threshold. On the opposite end, the combination of "other" mother + black father results in a drastic increase of the intercept term (=1733+821) above 2500 grams, the normal birthweight threshold.

## 6  Discussion
<!--
6. A discussion of whether your results can be interpretted causally.  In particular, include a discussion of what variables are not included in your analysis and the likely direction of omitted variable bias.  Also include a discussion of which included variables may bias your results by absorbing some of the causal effect of prenatal care.
-->

### 6.1 Causal Interpretation



### 6.2 Omitted Variable Bias
### 6.3 Included Variable Bias


## 7  Conclusion
<!--
7. A brief conclusion with a few high-level takeaways.
-->


<!-- Please limit all submissions to 30 pages.  Be sure to turn in both your pdf report and also your source code. -->